Commutative Character Sheaves on smooth group schemes over finite fields

### 2016-01-05 followup.tex
Cleaned up CCS section. Working on types.

\begin{abstract}
We describe a function-sheaf dictionary for one-dimensional characters of arbitrary smooth group schemes
over finite fields.  In previous work, we considered the case of smooth commutative group schemes and found
that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  In this paper,
we give a modification of the category of character sheaves that remedies this defect, and is also extensible to
non-commutative groups.  We use these \emph{commutative character sheaves} to geometrize types for
supercuspidal representations of $p$-adic groups. 
\end{abstract}

@@

Next, we broaden our scope further to encompass smooth group schemes $G$ that are not necessarily commutative.
In this paper, we do not consider Lusztig's geometrization of arbitrary characters of connected, reductive groups $G(k)$
using character sheaves \cite{lusztig:85a}*{Def.~2.10}, but focus on the case of $1$-dimensional characters
$G(k)^*$ assuming only that $G$ is smooth. The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more linear character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We define a category $\CCS(G)$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to G\ab(k)^*$
is an isomorphism.  Since $G\ab(k)^*$ surjects onto $G(k)^*$, for each character $\chi \in G(k)^*$ there is a commutative
character sheaf $\cs{L}$ on $G$ with $\TrFrob{G}(\cs{L}) = \chi$. Moreover, pullback along the quotient $q : G \to G\ab$
defines an equivalence of categories $\CCS(G\ab) \to \CCS(G)$.  Note that the functor $\CCS(G) \to \CS(G)$ is not
essentially surjective, missing the kinds of linear character sheaves highlighted by Kamgarpour.

@@

Suppose $G$ is commutative, so $G\der = 1$. 
Suppose $(\cs{L},\mu,\beta)$ is an object in $\CS'(G)$.
Then $\beta : \cs{L}_1\to \EE$ is an isomorphism in $\CS(1)$, and unique by \cite{cunningham-roe:13a}*{Thm 3.9}.
Tracing through the construction of $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ from $\beta : \cs{L}_1\to \EE$, we find \todo{I'm having some trouble seeing this.} that $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ is the canonical isomorphism coming from the equation $m = m \circ \theta$. 
Thus, when $G$ is commutative, Definition~\ref{def:CCS} recovers Definition~\ref{def:CCScom}.
The next result generalizes this observation.

\begin{theorem}\label{thm:Gab}
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}
%
Sections~\ref{ssec:equivariant1}, \ref{ssec:equivariant2}, \ref{ssec:descent} and \ref{ssec:closed} set up machinery needed to prove Theorem~\ref{thm:Gab}.
The precise definition of the functor $\CCS(G\ab) \to \CCS(G)$ will be given with the proof of Theorem~\ref{thm:Gab}, in Section~\ref{ssec:proof2}.

@@

Theorem~\ref{thm:Gab} shows that $\CCS(G)$ is a categorical solution to the problem that linear character sheaves on $G$ need not be trivial on $G\der$, as discussed at the beginning of Section~\ref{sec:noncom}; at the same time, it is the categorical solution to the problem, discussed in Section~\ref{ssec:defect}, that the dictionary between linear character sheaves on $G\ab$ and characters of $G\ab$, need not be perfect.
%
But a description of $\CCS(G)$ requires, at the very least, a description of isomorphism classes of objects, and maps. 
%

\begin{proposition}
The trace of Frobenius $\TrFrob{} : \CCSiso{G}\to G(\Fq)^*$ fits into the following diagram,
\[
\begin{tikzcd}
\ & & \CCSiso{G\ab} \arrow{d}{\TrFrob{}}[swap]{\iso} \arrow{r}{\iso} & \CCSiso{G} \arrow{d}{\TrFrob{}}& \\
1 \arrow{r} & \Delta^* \arrow{r} & G\ab(k)^* \arrow{r} & G(k)^* \arrow{r} & 1,
\end{tikzcd}
\]
where $\Delta$ denotes the image of the connecting homomorphism $G\ab(\Fq) \to \Hh^1(\Fq, G\der)$.
In particular, isomorphism classes of objects in $\CCS(G)$ are canonically identified with the group $G\ab(\Fq)^*$.
\end{proposition}

@@

\begin{proposition}
Every map in $\CCS(G)$ is either trivial or an isomorphism. 
For every commutative character sheaf on $G$, its automorphism group is
$
((\pi_0(G\ab))^*)^{\Frob{}}.
$
\end{proposition}

\begin{proof}
By Theorem~\ref{thm:Gab}, $\Aut_{\CCS(G)}(\cs{L},\mu,\beta) = \Aut_{\CCS(G\ab)}(\cs{L}\ab,\mu\ab)$. 
By \cite{cunningham-roe:13a}*{Thm 3.9},  $\Aut_{\CCS(G\ab)}(\cs{L}\ab,\mu\ab) = ((\pi_0(G\ab))^*)^{\Frob{}}$. \todo{Shouldn't it be coinvariants?}
\end{proof}

@@

@@

\begin{lemma}\label{lem:torsor}
If $q : G \to G_0$ is an $H$-torsor in the fppf topology 
then $q_H^*: \CS(G_0) \to \CS_H(G)$ is an equivalence.
\end{lemma}

\begin{proof}
First we recall that the comparison functor 
$\Loc(G_0) \to \Loc_H(G)$ is an equivalence.
To see this, observe first that $\Loc$ is a stack on schemes over $\Fq$ in the Zariski topology.
Then, observe that the comparison functor for Weil local systems is an equivalence for flat surjective morphisms of affine schemes over $\Fq$.
It follows that $\Loc$ is a stack over schemes over $\Fq$ in the fppf topology [Vistoli].
(This is a slight variation on the argument showing that quasicoherent sheaves are a stack over schemes in the fppf topology.)
Since $q : G\to G_0$ is an $H$-torsor in the fppf topology, is now follows from descent theory that the comparison functor $\Loc(G_0) \to \Loc_H(G)$ is an equivalence [Vistoli]. \todo{Need more specific references}

@@

%\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}] 
The abelianization $G\ab\ceq G/G\der$ exists in schemes over $\Fq$ \cite{demazure:SGA3-VIA}*{Thm. 3.2}.
Set $H=G\der$.
Let $j : H\hookrightarrow G$ be the inclusion and let $q : G \to G\ab$ be the quotient. 
With reference to \eqref{qH} and Section~\ref{ssec:closed}, consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(H) & \arrow{l}[swap]{j^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
 \CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} &  \CS_{H}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]
%
Here we use the notation $\CS'(G)$ from Section~\ref{ssec:noncomdef} and the observation that $\CS'(G)$ is precisely the category $\CS^{H}(G)$ from Section~\ref{ssec:closed} for $H=G\der$.
The forget
Since the quotient $q : G \to G\ab$ is a trivial $H$-torsor in the fppf topology \cite{demazure:SGA3-VIA}*{Thm. 3.2}, 
it now follows from Lemma~\ref{lem:torsor} that the comparison functor $\CS(G\ab)\to \CS_{H}(G)$ is an equivalence.
On the other hand, by Lemma~\ref{lem:beta}, $\CS'(G) \to \CS_{H}(G)$, defined by $(\cs{L},\mu,\beta)\mapsto (\cs{L},\mu, \nu)$, is an equivalence.
Since the essential image of the forgetful functor $\CS'(G)\to \CS(G)$ is precisely the category of $(\cs{L},\mu)\in \CS(G)$ such that $\cs{L}\vert_{H} \iso (\EE)_{H}$, this completes the proof of Proposition~\ref{prop:Gder_triv}.
%\end{proof}

@@

\section{Application to type theory for \texorpdfstring{$p$}{p}-adic groups}

@@

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from the data. 
Let us refer to the types arising in this manner as {\it Yu types}.
In this section we show that every Yu type is geometrizable.
% in a sense which will be made clear below.

A Yu type is a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, with certain properties. 
In Section~\ref{ssec:review} we recall properties of $(\,^\circ K,\,^\circ \rho)$ which we allow us, in Section~\ref{ssec:geometrization}, to find a pair $(G, \mathcal{F})$ where $G$ is a group scheme over $\Fq$ and $\mathcal{F}$ is a virtual constructible sheaf $\ell$-adic sheaf on $G$ from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

@@

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item[(0)] a compact group $\,^\circ K$;
\item[(1)] compact open subgroups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that \[\,^\circ K = J^0\cdots J^d;\]
\item[(2)] a continuous representation $\varpi_0$ of $J^0$, inflated from ...
\item[(3)] a quasicharacter $\varphi_i : J^0\cdots J^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
These representations $(\sigma_0, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties, which allow Yu to construct, for $i=0, \ldots, d-1$, further:
\begin{enumerate}
\item[(4)] a quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, and the structure of a $\Fq$-vector space on $V_{i+1}$, compatible with the abelian group structure above;
\item[(5)] a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$;
\item[(6)] an additive character $\psi_{i+1} : C_{i+1} \to \EEx$;
\item[(7)] a group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$;
\item[(8)] a quotient $J^{i+1} \to H^{i+1}$, where $H^{i+1}$ is a Heisenberg $p$-group;
\item[(9)] an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $(V_{i+1}, \langle\ ,\ \rangle_{i+1}, C_{i+1}, \psi_{i+1})$.
%\end{enumerate}
\item[(10)]The pair $(f_i, j_{i+1})$ is a symplectic action of $J^0\cdots J^i$ on $H^{i+1}$ in the sense of \cite{yu:construction}, so, together with the quotient $J^{i+1} \to H^{i+1}$, it defines a group homomorphism 
\begin{equation}\label{eq:symaction}
\begin{tikzcd}
h_{i+1} : (J^0\cdots J^i)\ltimes J^{i+1} \arrow{r} 
%& (J^0\cdots J^i)\ltimes H^{i+1} \arrow{r} 
& \Sp(V_{i+1})\ltimes V_{i+1}^\sharp.
\end{tikzcd}
\end{equation}
\end{enumerate}

From this information, Yu constructs 
a representation $\,^\circ \rho_i$ of $J^0\cdots J^i$, for each $i=0, \ldots, d$, such that $\,^\circ \rho_d = \,^\circ \rho$,
constructed as follows.
%\end{enumerate}
%We will need some more detail about the construction of the representations $\,^\circ \rho^i$, so we review it below.
%
We will need some detail about this construction,   so we review it here.

%Let $\varpi_1$ be the Heisenberg representation of the Heisenberg group $V_1^\sharp$, with central character $\psi_1$.
%Let $\varrho_1$ be the Weil representation of $\Sp(V_1)$ with the same representation space as $\omega_1$.
Let $\omega_{i+1}$ be the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
This determines the Weil representation $W_{i+1}$ on the same representation space as the Heisenberg representation of $V_{i+1}^\sharp$.
Let $W_{i+1} \ltimes \omega_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$.
\[
\begin{tikzcd}
\  & J^{i+1} \arrow{d} & \\
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp & \arrow{l}[swap]{h_{i+1}} (J^0 \cdots J^i) \ltimes J^{i+1} \arrow{r} \arrow{d} & J^0 \cdots J^i J^{i+1}\\
& J^0 \cdots J^i & 
\end{tikzcd}
\]

Now $h_1^*(W_1\ltimes \omega_1)$ is a representation of $J^0\ltimes J^1$, determined by the quasicharacter $\varphi_1$ of $J^1$. 
Write $\inf(\sigma_0)$ for the representation of $J^0 \ltimes J^1$ obtained by pulling back $\sigma_0$ along $J^0 \ltimes J^1 \to J^0$. 
By \cite{yu:01a}*{}, the representation 
$h_1^*(\omega_1) \otimes \varphi_1 \otimes \inf(\sigma_0)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 \ltimes 1) \cap (1 \ltimes J^1)$, so $h_1^*(\omega_1) \otimes \varphi_2 \otimes\inf(\sigma_0\otimes)$ descends to a representation $\,^\circ \rho_1$ of $J^0 J^1$.  
%
Likewise, the representation 
$h_{2}^*(\omega_{2}) \otimes \inf(\,^\circ \rho_1)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 J^1 \ltimes 1) \cap (1 \ltimes J^2)$, so $h_2^*(\omega_2) \otimes \inf(\,^\circ \rho_1)$ descends to a representation $\,^\circ \rho_2$ of $J^0 J^1 J^2$.  
%
Continuing in this way, one constructs the representation $\,^\circ \rho_d$ of $J^0\cdots J^d$. 
%
The character of $\,^\circ \rho_d$ on $\,^\circ K = J^0\cdots J^d$ satisfies
\[
\trace(\,^\circ \rho_d)(k_0 \cdots k_d) = \sigma_0(k_0) \left( \varphi_0(k_0) \cdots \varphi_d(k_d) \right) ,
\]
for $k_i\in J^i$.

>>>>

\begin{abstract}
We describe a function-sheaf dictionary for one-dimensional characters of arbitrary smooth group schemes
over finite fields.  In previous work, we considered the case of commutative smooth group schemes and found
that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  In this paper,
we give a modification of the category of character sheaves that remedies this defect, and is also extensible to
non-commutative groups.  
As an application, we use these \emph{commutative character sheaves} to geometrize types for
supercuspidal representations of $p$-adic groups. 
\end{abstract}

@@

Next, we broaden our scope further to encompass smooth group schemes $G$ that are not necessarily commutative.
In this paper, we do not consider Lusztig's geometrization of arbitrary characters of connected, reductive groups $G(k)$
using character sheaves \cite{lusztig:85a}*{Def.~2.10}, but focus on the case of $1$-dimensional characters
$G(k)^*$ assuming only that $G$ is smooth, allowing for the possibility that $G$ is not connected and not reductive.
The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more linear character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We define a category $\CCS(G)$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to G\ab(k)^*$
is an isomorphism.  
Since $G\ab(k)^*$ surjects onto $G(k)^*$, it follows that for each character $\chi \in G(k)^*$ there is a commutative
character sheaf $\cs{L}$ on $G$ with $\TrFrob{G}(\cs{L}) = \chi$. Moreover, we find that pullback along the quotient $q : G \to G\ab$
defines an equivalence of categories $\CCS(G\ab) \to \CCS(G)$.  Note that the functor $\CCS(G) \to \CS(G)$ is not
essentially surjective, missing the kinds of linear character sheaves highlighted by Kamgarpour.

@@

@@

Suppose $G$ is commutative, so $G\der = 1$. 
Suppose $(\cs{L},\mu,\beta)$ is an object in $\CS'(G)$.
Then $\beta : \cs{L}_1\to \EE$ is an isomorphism in $\CS(1)$, and unique by \cite{cunningham-roe:13a}*{Thm 3.9}.
Tracing through the construction of $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ from $\beta : \cs{L}_1\to \EE$, we find 
%\todo{I'm having some trouble seeing this.} 
that $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ is the canonical isomorphism coming from the equation $m = m \circ \theta$. 
Thus, when $G$ is commutative, Definition~\ref{def:CCS} recovers Definition~\ref{def:CCScom}.
The next result generalizes this observation.

\begin{theorem}\label{thm:Gab}
%Let $G$ be a smooth group scheme over a finite field $\Fq$. Let $\CCS(G)$ be the category of {\it commutative characters sheaves} on $G$; see Definition~\ref{def:CCS}.
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}
%
Sections~\ref{ssec:equivariant1}, \ref{ssec:equivariant2}, \ref{ssec:descent} and \ref{ssec:closed} set up machinery needed to prove Theorem~\ref{thm:Gab}.
The precise definition of the functor $\CCS(G\ab) \to \CCS(G)$ will be given with the proof of Theorem~\ref{thm:Gab}, in Section~\ref{ssec:proof2}.

Theorem~\ref{thm:Gab} shows that $\CCS(G)$ is a categorical solution to the problem that linear character sheaves on $G$ need not be trivial on $G\der$, as discussed at the beginning of Section~\ref{sec:noncom}; at the same time, it is the categorical solution to the problem, discussed in Section~\ref{ssec:defect}, that the dictionary between linear character sheaves on $G\ab$ and characters of $G\ab$, need not be perfect.
%
But a description of $\CCS(G)$ requires, at the very least, a description of isomorphism classes of objects, and maps.

\begin{corollary}\label{cor:Gab}
%Let $G$ be a smooth group scheme over a finite field $\Fq$. Let $\CCS(G)$ be the category of {\it commutative characters sheaves} on $G$; see Definition~\ref{def:CCS}.
Category $\CCS(G)$ is monoidal and $\CCSiso{G} \iso \Hom(G\ab(\Fq),\EEx)$, canonically.
Every map in $\CCS(G)$ is either trivial or an isomorphism, and the automorphism group of any object in $\CCS(G)$ is $\Hom(\pi_0(G\ab)_{\Frob{}},\EEx)$, canonically.
\end{corollary}

\subsection{Geometrization of characters} \label{ssec:geo}

\begin{theorem}\label{thm:geo}
%Let $G$ be a smooth group scheme over a finite field $\Fq$. Let $\CCS(G)$ be the category of {\it commutative characters sheaves} on $G$; see Definition~\ref{def:CCS}.
The trace of Frobenius $\TrFrob{} : \CCSiso{G}\to G(\Fq)^*$ fits into the following diagram,
\[
\begin{tikzcd}
\ & & \CCSiso{G\ab} \arrow{d}{\TrFrob{}}[swap]{\iso} \arrow{r}{\iso} & \CCSiso{G} \arrow{d}{\TrFrob{}}& \\
1 \arrow{r} & \Delta_G^* \arrow{r} & G\ab(k)^* \arrow{r} & G(k)^* \arrow{r} & 1,
\end{tikzcd}
\]
where $\Delta_G$ denotes the image of the connecting homomorphism $G\ab(\Fq) \to \Hh^1(\Fq, G\der)$. 
Thus, category $\CCS(G)$ geometrizes characters of $G(\Fq)$ in the following sense: for every group homomorphism $\chi : G(\Fq) \to \EEx$ there is an object $(\cs{L},\mu,\beta)$ in $\CCS(G)$ such that $\trFrob{\cs{L}} = \chi$. The geometrization of $\chi : G(\Fq) \to \EEx$ is not unique, but the group of isomorphism classes of possibilities are enumerated by $\Delta_G^*$.
\end{theorem}

@@

@@

%We close this section by gathering together the main result of this paper in one place.

@@

\begin{lemma}\label{lem:torsor}
If $q : G \to G_0$ is an $H$-torsor in the fpqc topology 
then $q_H^*: \CS(G_0) \to \CS_H(G)$ is an equivalence.
\end{lemma}

\begin{proof}
First we observe that the comparison functor 
$\Loc(G_0) \to \Loc_H(G)$ is an equivalence.
To see this, first recall that $\Loc$ is a stack on schemes over $\Fq$ in the Zariski topology.
Then, observe that the comparison functor for Weil local systems is an equivalence for flat surjective morphisms of affine schemes over $\Fq$.
It follows that $\Loc$ is a stack over schemes over $\Fq$ in the fpqc topology; see \cite{Vistoli:notes}*{Lem 4.25} for example.
(This is a slight variation on the argument showing that quasicoherent sheaves are a stack over schemes in the fpqc topology; see \cite{Vistoli:notes}*{Thm 4.23} for example.)
Since $q : G\to G_0$ is an $H$-torsor in the fpqc topology, is now follows from descent theory that the comparison functor $\Loc(G_0) \to \Loc_H(G)$ is an equivalence; see \cite{Vistoli:notes}*{Thm 4.46} for example.)
%\todo{Need more specific references}

@@

%\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}] 
By \cite{demazure:SGA3-VIA}*{Thm. 3.2}, the abelianization $G\ab\ceq G/G\der$ exists in schemes over $\Fq$ and the quotient $q: G \to G\ab$ is a $G\der$-torsor in the fppf topology.
%
To simplify notation below, set $H=G\der$
and let $j : H\hookrightarrow G$ be the inclusion. 
With reference to \eqref{qH} and Section~\ref{ssec:closed}, consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(H) & \arrow{l}[swap]{j^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
 \CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} &  \CS_{H}(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]
%
Here we use the notation $\CS'(G)$ from Section~\ref{ssec:noncomdef} and the observation that $\CS'(G)$ is precisely the category $\CS^{H}(G)$ from Section~\ref{ssec:closed} for $H=G\der$.
%
Since the quotient $q : G \to G\ab$ is an $H$-torsor in the fppf topology \cite{demazure:SGA3-VIA}*{Thm. 3.2}, 
it is also an $H$-torsor in the fpqc topology.
It now follows from Lemma~\ref{lem:torsor} that the comparison functor $\CS(G\ab)\to \CS_{H}(G)$ is an equivalence.
%
On the other hand, by Lemma~\ref{lem:beta}, $\CS'(G) \to \CS_{H}(G)$, defined by $(\cs{L},\mu,\beta)\mapsto (\cs{L},\mu, \nu)$, is an equivalence.
This completes the proof of Proposition~\ref{prop:Gder_triv}.
%\end{proof}

@@

\subsubsection{Proof of Corollary~\ref{cor:Gab}}\label{ssec:proof3}

Let us write $(\cs{L},\mu,\beta) \mapsto (\cs{L}\ab,\mu\ab)$ to indicate the equivalence appearing in Theorem~\ref{thm:Gab};
then \[\Aut_{\CCS(G)}(\cs{L},\mu,\beta) = \Aut_{\CCS(G\ab)}(\cs{L}\ab,\mu\ab).\] 
By \cite{cunningham-roe:13a}*{Thm 3.9},  $\Aut_{\CCS(G\ab)}(\cs{L}\ab,\mu\ab) = \Hom(\pi_0(G\ab)_{\Frob{}},\EEx)$.

\section{Application to type theory for \texorpdfstring{$p$}{p}-adic groups}\label{sec:types}

@@

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from Yu's data. 
Let us refer to the types arising in this manner as {\it Yu types}.
%In this section we show that every Yu type is geometrizable.
% in a sense which will be made clear below.
%
A Yu type is then a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, constructed in a certain way. 
%
Regarding the construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation.

\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}

In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to results push this observation further by geometrizing the representation $\,^\circ \rho$ of $\,^\circ K$: we find a group scheme $G$ over $\Fq$ such that $G(\Fq) = \,^\circ K$ and a virtual Weil sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\cs{F}} = \trace(\,^\circ \rho)$.
%from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

\subsection{Review of Jacobi theory over finite fields}

For use below, we recall some basic facts about the Heisenberg-Weil representation.

Let $V$ be a finite-dimensional vector space over a finite field $\Fq$ equipped with a symplectic paring $\langle\ ,\ \rangle : V\times V \to C$, where $C$ is a one-dimensional vector space over $\Fq$.
Then $V$ is even-dimensional and we may pick a `Lagrangian splitting' $V = U \oplus W$ such that $\langle v,v' \rangle = (u,w') - (u',w)$ for $v=(u,w)$ and $v'=(u',w')$.
Let $V^\sharp$ be the Heisenberg group determined by $(C, \langle\ ,\ \rangle)$.
Then $V^\sharp$ is a split extension of $V$ by $C$.
\[
\begin{tikzcd}
\ &&& U \arrow{d} &\\
1 \arrow{r} & C \arrow{r} & V^\sharp \arrow{r} & \arrow[bend right]{l} V \arrow[bend right]{u} \arrow[bend left]{d} \arrow{r} & 1\\
1 \arrow{r} & \arrow[equals]{u} C \arrow{r} & \arrow[bend right]{l}  A \arrow{u} \arrow{r} & \arrow[bend right]{l}  W \arrow{u} \arrow{r} & 1 
\end{tikzcd}
\]

Let $\psi : C \to \EEx$ be an additive character and let $\omega$ be the Heisenberg representation on $V^\sharp$ with central character $\psi : C \to \EEx$. 
Let $\Sp(V)$ be the symplectic group determined by the symplectic pairing $\langle\ ,\ \rangle$;
this group acts on $V^\sharp$.
Extend $\psi$ to $\psi_A : A \to \EEx$ in the obvious way. 
Then $\omega \ceq \operatorname{Ind}_A^{V^\sharp}(\psi_A)$ is called the Heisenberg representation of $V^\sharp$.

Observe that $\Sp(V)$ acts on $V^\sharp$.
The Heisenberg representation determines a representation $\pi_{\omega}$ of $\Sp(V)$ with the same representation space as $\omega$ and with the defining property: for each $g\in \Sp(V)$, $\pi_\omega(g)$ determines an isomorphism of representations $\omega^g \to \omega$, through the Egorov identity \cite{Gurevich-Hadani:geometric_Weil}*{(1.3.1)}.
Let $W = \pi_\omega \ltimes \omega$ be the representation of the Jacobi group $\Sp(V)\ltimes V^\sharp$ given by $\omega$ and $\pi_\omega$.

@@

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item[(0)] a sequence of compact groups 
\[
\,^\circ K^0 \subseteq \,^\circ K^1 \subseteq \cdots \subseteq \,^\circ K^d= \,^\circ K;
\]
\item[(1)] a continuous representation $\varrho$ of $\,^\circ K^0$;
\item[(2)] a quasicharacter $\varphi_i : \,^\circ K^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
The representations $(\varrho, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties which allow Yu to construct representations $\,^\circ\rho_i$ of $\,^\circ K^i$, for $i=0, \ldots, d-1$, such that $\,^\circ\rho_0 = \varrho$ and $\,^\circ\rho_d = \,^\circ\rho$.

Since we will need some facts about this construction, we review some further detail here.
First, Yu introduces compact groups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that 
\[
\,^\circ K^i = J^0\cdots J^{i}
\]
and a natural action of $\,^\circ K^i$ on $J^{i+1}$, for $i=0, \ldots d-1$, defining the groups
\[
\,^\circ K^i \ltimes J^{i+1}.
\]
He then introduces the following quotients of $J^{i+1}$ and $\,^\circ K^i$, for $i=0, \ldots d-1$.
\begin{enumerate}
\item[(Q.1)] A quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, the latter also given the structure of a $\Fq$-vector space.
\end{enumerate}
The vector space $V_{i+1}$ is then equipped with a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$, itself equipped with an additive character $\psi_{i+1} : C_{i+1} \to \EEx$.
\begin{enumerate}
\item[(Q.2)] A quotient $J^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $V_{i+1}$, $C_{i+1}$, $\langle\ ,\ \rangle_{i+1}$ and $\psi_{i+1}$.
\end{enumerate}
In fact, the quotient $J^{i+1} \to V_{i+1}^\sharp$ factors through a quotient $J^{i+1} \to H^{i+1}$ and an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $H^{i+1}$ is a Heisenberg $p$-group in the sense of \cite{yu:construction}*{}.
\begin{enumerate}
\item[(Q.3)] A quotient $\,^\circ f_i : \,^\circ K^i \to \Sp(V_{i+1})$, such that the pair $(f_i, j_{i+1})$ is a symplectic action of $\,^\circ K^i$ on $H^{i+1}$ in the sense of \cite{yu:construction}.
\end{enumerate}
Taken together, these quotients define another quotient:
\begin{enumerate}
\item[(Q)] A quotient $h_{i+1} : \,^\circ K^i \ltimes J^{i+1} \to  \Sp(V_{i+1})\ltimes V_{i+1}^\sharp$ making the following diagram commute.
\[
\begin{tikzcd}
1 \arrow{r} & J^{i+1} \arrow{d} \arrow{r} & \,^\circ K^i \ltimes J^{i+1} \arrow{d}{h_{i+1}} \arrow{r} & \,^\circ K^i \arrow{r} \arrow{d}{\,^\circ f_i} & 1\\ 
1 \arrow{r} & V_{i+1}^\sharp \arrow{r} & \Sp(V_{i+1}) \ltimes V_{i+1}^\sharp \arrow{r} & \Sp(V_{i+1}) \arrow{r} & 1
\end{tikzcd}
\]
\end{enumerate}

The representations $\,^\circ \rho_i$ are defined recursively. Set $\,^\circ \rho_0 = \varrho$.
Suppose $i=0, \ldots, d-1$.
The key to defining the representation $\,^\circ\rho_{i+1}$ from $\,^\circ \rho_i$ is the following diagram.
\[
\begin{tikzcd}
\ && \,^\circ K^{i}\cap J^{i+1}\arrow{d} && \\
1 \arrow{r} & J^{i+1} \arrow{r} & \,^\circ K^i \ltimes J^{i+1} \arrow{d} \arrow{r} & \,^\circ K^i \arrow{r} & 1\\
&& \,^\circ K^{i+1} &&
\end{tikzcd}
\]
Let $\omega_{i+1}$ be the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
This determines a Weil representation on the same representation space as $\omega_{i+1}$.
Let $W_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$.
Then $h_{i+1}^*(W_{i+1})$ is a representation of $\,^\circ K^i \ltimes J^{i+1}$.
Write $\inf(\,^\circ\rho_i \otimes \varphi_i)$ for the representation of $\,^\circ K^i \ltimes J^{i+1}$ obtained by pulling back $\,^\circ\rho_i$ along $\,^\circ K^i \ltimes J^{i+1} \to \,^\circ K^i$. 
By \cite{yu:01a}*{}, the representation 
\[
h_{i+1}^*(W_{i+1}) \otimes \inf(\,^\circ\rho_i \otimes \varphi_i)
\] 
of $\,^\circ K^i \ltimes J^{i+1}$ is trivial on $\,^\circ K^{i}\cap J^{i+1}$, so $h_{i+1}^*(W_{i+1}) \otimes \inf(\,^\circ\rho_i \otimes \varphi_i)$ descends to $\,^\circ K^{i+1}$. This defines $\,^\circ \rho_{i+1}$.\todo{I didn't use $\varphi_d$, so this can't be quite right.}

%The character of $\,^\circ \rho_d$ on $\,^\circ K^d = \,^\circ K = J^0 \cdots J^d$ satisfies
%\[
%\trace(\,^\circ \rho)(k_0 \cdots k_d) = \varrho(k_0)  \varphi_0(k_0) \cdots \varphi_d(k_d),
%\]
%for $k_i\in J^i$.
